Top-down versus bottom-up cohesiveness in microbial community coalescence

### 2021-05-25 draft/draft_v3.tex
minor changes to draft

\begin{equation}
\frac{dR_\alpha}{dt} = 
- \sum_j N_j c_{j\alpha} R_\alpha
+ \sum_j \sum_\beta N_j c_{j\beta} R_\beta
\left[
l_\beta D_{j\alpha\beta} \frac{w_\beta}{w_\alpha}
\right]
\label{eq:crm-r}
\end{equation}
%
These equations can take slightly different forms in certain cases,
e.g. if the primary resource is supplied continuously instead of at the beggining
of each growth cycle \cite{Marsland2019,Marsland2020},
but they are a good approximation for the community dynamics between consecutive
serial dilutions in our setup.
Here, we assembled \textit{in silico} communities by randomly
sampling a set of species from a pool,
then integrating equations \ref{eq:crm-n} and \ref{eq:crm-r},
diluting the final abundances,
replenishing the primary resource (by default denoted as $R_0$),
and repeating the process until generational equilibrium was achieved
(\methodsref{sim}).
Coalescence simulations were carried out following the same logic,
this time seeding the coalesced communities by mixing the invasive and resident
communities instead of sampling from a species pool.

@@

\figref[b]{fig3} shows that, in the simulations,
many dominants could not invade on their own
(or could only do so at very low final relative abundances, below 0.1)
but were able to reach high frequencies when they were accompanied by their cohorts.
This indicates that positive bottom-up co-selection is frequent and potentially very strong,
while negative bottom-up co-selection is far more uncommon.
We then asked whether the ability of the pairwise competition of dominants
to predict coalescence outcomes was dependent on the strength of bottom-up co-selection.
We divided our simulations into two subsets:
the first one was comprised of the instances where positive bottom-up co-selection was strong
(i.e. dots in the green shaded region of \figref[b]{fig3}),
the second set included all other cases
(dots near the diagonal of \figref[b]{fig3}).
We reexamined our original simulations and plotted the frequency of the invasive dominant
in pairwise competition with the resident dominant versus the relative similarity between
the invasive and coalesced communities,
i.e. the same plot as in \figref[b]{fig2}, for each subset.
We found that when bottom-up positive co-selection is strong,
the pairwise competition of dominants is not predictive of coalescence outcomes
(\figref[c]{fig3}, left panel) and vice-versa (\figref[c]{fig3}, right panel).

>>>>

\begin{equation}
\frac{dR_\alpha}{dt} = 
- \sum_j N_j c_{j\alpha} R_\alpha
+ \sum_j \sum_\beta N_j c_{j\beta} R_\beta
\left[
l_\beta D_{j\alpha\beta} \frac{w_\beta}{w_\alpha}
\right]
\label{eq:crm-r}
\end{equation}
%
These equations can take slightly different forms in certain cases,
e.g. if the primary resource is supplied continuously instead of at the beggining
of each growth cycle \cite{Marsland2019,Marsland2020},
but they are a good approximation for the community dynamics between consecutive
serial dilutions in our setup.
Here, we assembled \textit{in silico} communities by randomly
sampling a set of species from a pool,
then integrating equations \ref{eq:crm-n} and \ref{eq:crm-r},
diluting the final abundances,
replenishing the primary resource,
and repeating the process until generational equilibrium was achieved
(\methodsref{sim}).
Coalescence simulations were carried out following the same logic,
this time seeding the coalesced communities by mixing the invasive and resident
ones instead of sampling from a species pool.

@@

\figref[b]{fig3} shows that, in the simulations,
many dominants could not invade on their own
(or could only do so at very low final relative abundances, below 0.1)
but were able to reach high frequencies when they were accompanied by their cohorts.
This indicates that positive bottom-up co-selection is frequent and potentially very strong,
while negative bottom-up co-selection is far more uncommon.
We then asked whether the ability of the pairwise competition of dominants
to predict coalescence outcomes was dependent on the strength of bottom-up co-selection.
We divided our simulations into two subsets:
the first one was comprised of the instances where positive bottom-up co-selection was strong
(i.e. dots in the green shaded region of \figref[b]{fig3}),
the second set included all other cases
(dots near the diagonal of \figref[b]{fig3}).
We reexamined our original simulations and plotted the frequency of the invasive dominant
in pairwise competition
\iffalse
with the resident dominant
\fi
versus the relative similarity between
the invasive and coalesced communities,
i.e. the same plot as in \figref[b]{fig2}, for each subset.
We found that when bottom-up positive co-selection is strong,
the pairwise competition of dominants is not predictive of coalescence outcomes
(\figref[c]{fig3}, left panel) and vice-versa (\figref[c]{fig3}, right panel).
